PyQBench: a Python library for benchmarking gate-based quantum computers

### 2023-10-09 txt/manuscript_short.tex
Explain NISQ devices

Noisy Intermediate-Scale Quantum (NISQ)~\cite{preskill} devices are storming the market, with a wide
selection of devices based on different architectures and accompanying software solutions. Among
hardware providers offering public access to their gate--based devices, one could mention Rigetti
\cite{rigetti}, IBM \cite{ibmq}, Oxford Quantum Group \cite{oxforf}, IonQ \cite{ionq} or Xanadu
\cite{xanadu}. Other vendors offer devices operating in different paradigms. Notably, one could
mention D-Wave \cite{dwave} and their quantum annealers, or QuEra devices \cite{quera} based on
neural atoms. Most vendors provide their own software stack and application programming interface
for accessing their devices. To name a few, Rigetti's computers are available through their Forest
SDK \cite{sdk} and PyQuil library \cite{pyquil} and IBM Q \cite{ibmq} computers can be accessed
through Qiskit \cite{qiskit} or IBM Quantum Experience web interface \cite{ibmqplatform}. Some cloud
services, like Amazon Braket \cite{amazon}, offer access to several quantum devices under a unified
API. On top of that, several libraries and frameworks can integrate with multiple hardware vendors.
Examples of such frameworks include IBM Q's Qiskit, Zapata Computing's Orquestra \cite{zapata}, XACC \cite{xacc} and NVIDIA's CUDA Quantum \cite{cudaquantum}.

It is well known that NISQ devices have their limitations \cite{preskillnew}. The question is to
what extent those devices can perform meaningful computations? To answer this question, one has to
devise a methodology for benchmarking them. For gate--based computers, on which this paper focuses,
there already exist several approaches. One could mention randomized benchmarking
\cite{liu2022sampling, knill2008randomized, wallman2014randomized, helsen2022general,
cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating,
moll2018quantum, pelofske2022volume}.

>>>>

Quantum devices are storming the market, with a wide
selection of devices based on different architectures and accompanying software solutions. Among
hardware providers offering public access to their gate--based devices, one could mention Rigetti
\cite{rigetti}, IBM \cite{ibmq}, Oxford Quantum Group \cite{oxforf}, IonQ \cite{ionq} or Xanadu
\cite{xanadu}. Other vendors offer devices operating in different paradigms. Notably, one could
mention D-Wave \cite{dwave} and their quantum annealers, or QuEra devices \cite{quera} based on
neural atoms. Most vendors provide their own software stack and application programming interface
for accessing their devices. To name a few, Rigetti's computers are available through their Forest
SDK \cite{sdk} and PyQuil library \cite{pyquil} and IBM Q \cite{ibmq} computers can be accessed
through Qiskit \cite{qiskit} or IBM Quantum Experience web interface \cite{ibmqplatform}. Some cloud
services, like Amazon Braket \cite{amazon}, offer access to several quantum devices under a unified
API. On top of that, several libraries and frameworks can integrate with multiple hardware vendors.
Examples of such frameworks include IBM Q's Qiskit, Zapata Computing's Orquestra \cite{zapata}, XACC \cite{xacc} and NVIDIA's CUDA Quantum \cite{cudaquantum}.

All of the currently available quantum computers have to be classified as NISQ, or Noisy Intermediate-Scale
Quantum devices. The term NISQ denotes quantum computers of an intermediate size expected in the next few years, typically comprising 50 to a few hundred qubits. The significance of reaching 50 qubits lies in surpassing the simulation capabilities of current digital supercomputers. The term 'Noisy' underscores the imperfect control over qubits, with noise imposing substantial limitations on the achievable outcomes of quantum devices in the short term \cite{preskill,preskillnew}. The question is to what extent those devices can perform meaningful computations? To answer this question, one has to
devise a methodology for benchmarking them. For gate--based computers, on which this paper focuses,
there already exist several approaches. One could mention randomized benchmarking
\cite{liu2022sampling, knill2008randomized, wallman2014randomized, helsen2022general,
cornelissen2021scalable}, benchmarks based on the quantum volume \cite{cross2019validating,
moll2018quantum, pelofske2022volume}.
